### 2016-03-21 manuscript.Rnw
fix typo, reflow paragraphs

The full Open Tree currently does not have any branch lengths associated with
it, therefore parametric comparative methods cannot be used directly on the
subtrees returned by OTL (although the OTL treestore contains the raw
published source phylogenies, complete with branch lengths and node
annotations; see below). However, resources and methods are being developed to add
branch lengths to these topological subtrees \citep[e.g.,][]{Ksepka2015} or use
topological trees to identify phylogenetically equivalent species to increase
overlap between chronograms and species trait data \citep{Pennell2015}. Without
branch lengths, these subtrees are nonetheless useful to illustrate relationships
among species, or to map traits on a phylogeny.

@@

The package is well-documented, and includes three package
vignettes (documents that demonstrate the use of the package and contain executable
R code). There is also an extensive test-suite that covers both the internal
functions that \texttt{rotl} uses to connect to OTL and public functions that users
apply to access and process data.

@@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{phylo}) associated with a particular
study. Metadata about the study (e.g., citation information, information about
the curator for the study, and other technical information regarding the import
of this study) can be obtained
using the function \texttt{get\_study\_meta}.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data
are made available online and accessible using APIs, several packages have been
developed to interact and download these datasets directly in R, thereby
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general. For instance, the
rOpenSci-developed package \texttt{TreeBase} \citep{Boettiger2012TB} allows users
to access phylogenies stored in treeBASE (\url{www.treebase.org}).
\texttt{rotl} contributes to this initiative, and greatly extends the number of
taxa for which phylogenetic data can be retrieved within R, while allowing the
data from OTL to be combined with other sources easily.

>>>>

Currently, the synthetic tree does not have any branch lengths associated with
it, therefore parametric comparative methods cannot be used directly on the
subtrees returned by OTL (although the OTL treestore contains the raw published
source phylogenies, complete with branch lengths and node annotations; see
below). However, resources and methods are being developed to add branch lengths
to these topological subtrees \citep[e.g.,][]{Ksepka2015} or use topological
trees to identify phylogenetically equivalent species to increase overlap
between chronograms and species trait data \citep{Pennell2015}. Without branch
lengths, these subtrees are nonetheless useful to illustrate relationships among
species, or to map traits on a phylogeny.

@@

The package is well-documented, and includes three package vignettes (documents
that demonstrate the use of the package and contain executable R code). There is
also an extensive test-suite that covers both the internal functions that
\texttt{rotl} uses to connect to OTL, and public functions that users apply to
access and process data.

@@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{phylo} objects) associated with a particular study. Metadata
about the study (e.g., citation information, information about the curator for
the study, and other technical information regarding the import of this study)
can be obtained using the function \texttt{get\_study\_meta}.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data
are made available online and accessible using APIs, several packages have been
developed to interact and download these datasets directly in R, thereby
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general \citep{Boettiger2015a}. For
instance, the rOpenSci-developed package \texttt{TreeBase}
\citep{Boettiger2012TB} allows users to access phylogenies stored in treeBASE
(\url{www.treebase.org}).  \texttt{rotl} contributes to this initiative, and
greatly extends the number of taxa for which phylogenetic data can be retrieved
within R, while allowing the data from OTL to be combined with other sources
easily.
